---
title: "Flu/SC2 & N1/N2 assay Ct Comparison"
author: "Vivian Hogan"
date: "`r Sys.Date()`"
output: html_document
---
<br>

##### <span style="color: gray;">For the most part N1 and N2 Cts are quite comparable to the Ct values obtained in the Flu/Sc2 assay. Important exceptions (from a total of 176 specimens): </span>
<br>

##### <span style="color: gray;">Specimens can have a high Ct on one assay and be negative/inconclusive on the other </span>
##### <span style="color: gray;">Two specimens became positive on the N1/N2 assay with Cts around 30 after being negative on Flu/Sc2. </span>
##### <span style="color: blue;">Check how baseline determination effects these results - Does not seem to effect discrepancies. Auto baseline seems reasonable on Flu/Sc2 assay. But shape of some curves under threshold is weird.</span>
<br>

```{r setup, include=FALSE}

#LOAD  LIBRARIES
library(knitr)
library(tidyverse)
library(dplyr)  
library(readr)
library(viridis)
library(plotly)
library(ggplot2)
library(ggpubr)
library(readxl)
library(purrr)
library(DT)
library(RColorBrewer)
knitr::opts_chunk$set(echo = FALSE)
```

```{r,echo=FALSE, message=FALSE, warning=FALSE}
#import data
data26 <- read_excel(path = "/Users/vivian/EPICC/FluSc2_Comp/20220801_wk26_N1_N2_comparison.xlsx", sheet = "Sheet8")%>%
  mutate(week = "Week 26") %>%
  mutate(SC2 = Sc2) %>%
  select(-Sc2)


data <- read_excel(path = "/Users/vivian/EPICC/FluSc2_Comp/20220627_wk24-001-088_N1_N2_comparison.xlsx", sheet = "Sheet8") %>%
  mutate(week = "Week 24") %>%
  bind_rows(data26) %>%
  arrange(week)


```

```{r,echo=FALSE, message=FALSE, warning=FALSE}
#add columns with percent difference
data <- data %>%
  mutate(N1_diff = (SC2 - N1)/SC2 * 100) %>%
  mutate(N2_diff = (SC2 - N2)/SC2 * 100)

#table with N1 diff > 15%
N1_outliers <- data %>%
  filter(N1_diff > 15 | N1_diff < -15) %>%
  mutate(N1_diff = signif(N1_diff, 3))

#table with N2 diff > 15%
N2_outliers <- data %>%
  filter(N2_diff > 15 | N2_diff < -15) %>%
  mutate(N2_diff = signif(N2_diff, 3))

```

```{r,echo=FALSE, message=FALSE, warning=FALSE}
#Scatter plot, outline outliers

N1 <- data %>%
  plot_ly(x=~SC2, y =~N1) %>%
            add_trace(
            type = 'scatter', 
            mode = 'markers',
            color = data$week, 
            colors = c("deeppink4", "deeppink1"),
            text = ~paste('</br>', SampleID,
                          '</br>Week: ', week,
                        '</br>Flu SC2 Assay: ', SC2, 
                        '</br>N1: ', N1,
                        '</br>N2: ', N2,
                        '</br>N1 diff: ', N1_diff), 
            hoverinfo = 'text')  %>%
  
  add_trace(x=N1_outliers$SC2, y = N1_outliers$N1,
            name = "N1 Outliers",
            type = 'scatter', 
            mode = 'markers', 
            marker = list(color = "blue", symbol = 'star-diamond-open'),
            text = ~paste('</br>', N1_outliers$SampleID,
                          '</br>Week: ', N1_outliers$week,
                        '</br>Flu SC2 Assay: ', N1_outliers$SC2, 
                        '</br>N1: ', N1_outliers$N1,
                        '</br>N2: ', N1_outliers$N2,
                        '</br>N1 diff: ', N1_outliers$N1_diff), 
            hoverinfo = 'text') %>%
  
   layout(title ='N1 Ct Comparison',
         autosize = F, width = 650, height = 500,
         yaxis = list(title = 'N1 Ct', 
                      range=list(0,45),
                      tickvals = seq(0, 45, 5)),
         xaxis = list(title = 'Flu Sc2 Assay Ct', 
                      range=list(0,45), 
                      tickvals = seq(0, 45, 5)))
N1
```  
<br>

```{r,echo=FALSE, message=FALSE, warning=FALSE}  
N2 <- data %>%
  plot_ly(x=~SC2, y =~N2) %>%
  add_trace(type = 'scatter', 
            mode = 'markers', 
            color = data$week,
            colors = c("deeppink4", "deeppink1"),
            text = ~paste('</br>', SampleID,
                          '</br>Week: ', week,
                        '</br>Flu SC2 Assay: ', SC2, 
                        '</br>N1: ', N1,
                        '</br>N2: ', N2,
                        '</br>N2 diff: ', N2_diff), 
            hoverinfo = 'text')  %>%
    
    add_trace(x=N2_outliers$SC2, y = N2_outliers$N2,
            name = "N2 Outliers",
            type = 'scatter', 
            mode = 'markers', 
            marker = list(color = "blue", symbol = 'star-diamond-open'),
            text = ~paste('</br>', N2_outliers$SampleID,
                          '</br>Week: ', N2_outliers$week,
                        '</br>Flu SC2 Assay: ', N2_outliers$SC2, 
                        '</br>N1: ', N2_outliers$N1,
                        '</br>N2: ', N2_outliers$N2,
                        '</br>N2 diff: ', N2_outliers$N2_diff), 
            hoverinfo = 'text') %>%
  
    
    layout(title ='N2 Ct Comparison',
         autosize = F, width = 650, height = 500,
         yaxis = list(title = 'N2 Ct', 
                      range=list(0,45),
                      tickvals = seq(0, 45, 5)),
         xaxis = list(title = 'Flu Sc2 Assay Ct', 
                      range=list(0,45), 
                      tickvals = seq(0, 45, 5)))

N2
  
```
