---
title: "Midnight vs Artic"
author: "Vivian Hogan"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(stringr)
library(readxl)
library(plotly)
library(viridis)
library(DT)
library(reshape2)
library(cowplot)
library(lubridate)
library(RColorBrewer)

knitr::opts_chunk$set(echo = FALSE)
```
```{r}
#load data
performance <- read_excel("/Users/vhogan/Documents/ARIA/Midnight/231004_NB2_Artic_performance.xlsx", sheet = "midnight")  %>%
  rename(SampleID = Sample) %>%
  mutate(`%>=20x` = as.numeric(`%>=20x`))

nextclade <- read_excel("/Users/vhogan/Documents/ARIA/Midnight/231004_NB2_Artic_lineage.xlsx", sheet = "Nextclade-midnight") %>%
  rename(SampleID = seqName) %>%
  mutate(qcoverallScore = as.numeric(qcoverallScore))

pangolin <- read_excel("/Users/vhogan/Documents/ARIA/Midnight/231004_NB2_Artic_lineage.xlsx", sheet = "Pangolin-midnight")  %>%
  rename(SampleID = taxon)

lib_files <- read_excel("/Users/vhogan/Documents/ARIA/Midnight/20231003_wk39-Aria_v5-midv2-InfA-RSVB-ADVhex_KL.xlsx", 'Library Quant (C)', skip=40) %>%
  rename(SampleID = `Sample ID`) %>%
  rename(lib_quant = 5) %>%
  filter(SampleID!=0) %>%
  filter(str_detect(SampleID, "^A")) %>%
  select(SampleID, lib_quant)
  

amp_files <- read_excel(path = "/Users/vhogan/Documents/ARIA/Midnight/20231003_wk39-Aria_v5-midv2-InfA-RSVB-ADVhex_KL.xlsx", sheet = 4, skip=39) %>%
  rename(SampleID = `Sample ID`) %>%
  rename(amp_quant = 5) %>%
  filter(SampleID!=0) %>%
  filter(str_detect(SampleID, "^A")) %>%
  select(SampleID, amp_quant)


#join data, add discrete grouping of RP Cts, rename test_type
#convert to date
joined_data <- performance %>%
  left_join(nextclade, by = "SampleID") %>%
  left_join(pangolin, by = "SampleID") %>%
  left_join(amp_files, by = "SampleID") %>%
  left_join(lib_files, by = "SampleID") 

```

### <span style="color: blue;">Comparison of Amp Quant for same sample with Midnight vs Artic</span>

```{r, warning = FALSE}
#Use melt function to create row for each subject/primer set with Amp Quant
amp_data_melt <- melt(joined_data, id.vars='SampleID', measure.vars=c('amp_quant')) %>%
                 separate(SampleID, into = c("Sample", "Primer"), sep = "_") %>%
                 mutate(Primer = case_when(Primer == "midV2" ~ "Midnight",
                                            Primer == "v5" ~ "Artic"))

#Using variable (test type) and value (Rp Ct) to create connected box plot with lines between subject ID
amp_plot <- amp_data_melt %>%
    ggplot(aes(x = Primer, y = value, fill = Primer)) + 
    geom_boxplot(alpha = 0.1) +
    geom_line(aes(group = Sample), linewidth=.6, position = position_dodge(0.5), color = "gray") + #line between same SampleID
    geom_point(aes(group = Sample), size=2, shape=21, position = position_dodge(0.5)) + #point for each sample
    labs(#can also add title, subtitle and/or caption
          title = "Amp Quant Comparison",
          y = "Amp Quant",
          x = "Primer Set") +
    theme_cowplot() +
    scale_fill_manual(values = c("orangered", "slateblue")) +
    theme(legend.position = "none") #+ #remove legend
    #ylim(20,37.418) #set limits of y axis

amp_plot
```

### <span style="color: blue;">Comparison of Lib Quant for same sample with Midnight vs Artic</span>

```{r, warning = FALSE}
#Use melt function to create row for each subject/primer set with Amp Quant
lib_data_melt <- melt(joined_data, id.vars='SampleID', measure.vars=c('lib_quant')) %>%
                 separate(SampleID, into = c("Sample", "Primer"), sep = "_") %>%
                 mutate(Primer = case_when(Primer == "midV2" ~ "Midnight",
                                            Primer == "v5" ~ "Artic"))

#Using variable (test type) and value (Rp Ct) to create connected box plot with lines between subject ID
lib_plot <- lib_data_melt %>%
    ggplot(aes(x = Primer, y = value, fill = Primer)) + 
    geom_boxplot(alpha = 0.1) +
    geom_line(aes(group = Sample), linewidth=.6, position = position_dodge(0.5), color = "gray") + #line between same SampleID
    geom_point(aes(group = Sample), size=2, shape=21, position = position_dodge(0.5)) + #point for each sample
    labs(#can also add title, subtitle and/or caption
          title = "Library Quant Comparison",
          y = "Library Quant",
          x = "Primer Set") +
    theme_cowplot() +
    scale_fill_manual(values = c("orangered", "slateblue")) +
    theme(legend.position = "none") #+ #remove legend
    #ylim(20,37.418) #set limits of y axis

lib_plot
```

### <span style="color: blue;">Comparison of Percent Genome Coverage at 20x for same sample with Midnight vs Artic</span>

```{r, warning = FALSE}
#Use melt function to create row for each subject/primer set with Amp Quant
cov_melt <- melt(joined_data, id.vars='SampleID', measure.vars=c('%>=20x')) %>%
                 separate(SampleID, into = c("Sample", "Primer"), sep = "_") %>%
                 mutate(Primer = case_when(Primer == "midV2" ~ "Midnight",
                                            Primer == "v5" ~ "Artic"))

#Using variable (test type) and value (Rp Ct) to create connected box plot with lines between subject ID
cov_plot <- cov_melt %>%
    ggplot(aes(x = Primer, y = value, fill = Primer)) + 
    #geom_boxplot() +
    geom_line(aes(group = Sample), linewidth=.6, position = position_dodge(0.5), color = "gray") + #line between same SampleID
    geom_point(aes(group = Sample), size=2, shape=21, position = position_dodge(0.5)) + #point for each sample
    labs(#can also add title, subtitle and/or caption
          title = "Coverage Comparison (>100 = bad)",
          y = "Coverage",
          x = "Primer Set") +
    theme_cowplot() +
    scale_fill_manual(values = c("orangered", "slateblue")) +
    theme(legend.position = "none") #+ #remove legend
    #ylim(20,37.418) #set limits of y axis

cov_plot
```
### <span style="color: blue;">Comparison of QC Overall Score for same sample with Midnight vs Artic</span>

```{r, warning = FALSE}
#Use melt function to create row for each subject/primer set with Amp Quant
qc_data_melt <- melt(joined_data, id.vars='SampleID', measure.vars=c('qcoverallScore')) %>%
                 separate(SampleID, into = c("Sample", "Primer"), sep = "_") %>%
                 mutate(Primer = case_when(Primer == "midV2" ~ "Midnight",
                                            Primer == "v5" ~ "Artic"))

#Using variable (test type) and value (Rp Ct) to create connected box plot with lines between subject ID
qc_plot <- qc_data_melt %>%
    ggplot(aes(x = Primer, y = value, fill = Primer)) + 
    #geom_boxplot() +
    geom_line(aes(group = Sample), linewidth=.6, position = position_dodge(0.5), color = "gray") + #line between same SampleID
    geom_point(aes(group = Sample), size=2, shape=21, position = position_dodge(0.5)) + #point for each sample
    labs(#can also add title, subtitle and/or caption
          title = "QC Overall Score Comparison (>100 = bad)",
          y = "QC Overall Score",
          x = "Primer Set") +
    theme_cowplot() +
    scale_fill_manual(values = c("orangered", "slateblue")) +
    theme(legend.position = "none") #+ #remove legend
    #ylim(20,37.418) #set limits of y axis

qc_plot
```

### <span style="color: blue;">Comparison of Total Missing for same sample with Midnight vs Artic</span>

```{r, warning = FALSE}
#Use melt function to create row for each subject/primer set with Amp Quant
missing_melt <- melt(joined_data, id.vars='SampleID', measure.vars=c('totalMissing')) %>%
                 separate(SampleID, into = c("Sample", "Primer"), sep = "_") %>%
                 mutate(Primer = case_when(Primer == "midV2" ~ "Midnight",
                                            Primer == "v5" ~ "Artic"))

#Using variable (test type) and value (Rp Ct) to create connected box plot with lines between subject ID
missing_plot <- qc_data_melt %>%
    ggplot(aes(x = Primer, y = value, fill = Primer)) + 
    #geom_boxplot() +
    geom_line(aes(group = Sample), linewidth=.6, position = position_dodge(0.5), color = "gray") + #line between same SampleID
    geom_point(aes(group = Sample), size=2, shape=21, position = position_dodge(0.5)) + #point for each sample
    labs(#can also add title, subtitle and/or caption
          title = "Total Missing Comparison (>100 = bad)",
          y = "Total Missing Score",
          x = "Primer Set") +
    theme_cowplot() +
    scale_fill_manual(values = c("orangered", "slateblue")) +
    theme(legend.position = "none") #+ #remove legend
    #ylim(20,37.418) #set limits of y axis

missing_plot
```

### <span style="color: blue;">Nextclade Comparison</span>
##### <span style="color: grey;">All match except A10-08514-4-Ag and A10-08549-6-Ag are able to make more specific calls with Artic primers</span>

```{r}
nextclade <- joined_data %>%
  select(SampleID, `%>=20x`, qcoverallScore, totalMissing, partiallyAliased) %>%
  separate(SampleID, into = c("SampleID", "Primer"), sep = "_") %>%
  mutate(Primer = case_when(Primer == "midV2" ~ "Midnight",
                            Primer == "v5" ~ "Artic")) %>%
  pivot_wider(names_from = c(Primer), values_from = partiallyAliased)

datatable(nextclade, options = list(pageLength = -1)) #pageLength, showing all
```

### <span style="color: blue;">Pangolin Comparison</span>
##### <span style="color: grey;">All match except A10-08549-6-Ag is unable to make a call with Midnight primers</span>

```{r}
pangolin <- joined_data %>%
  select(SampleID, `%>=20x`, qcoverallScore, totalMissing, lineage) %>%
  separate(SampleID, into = c("Sample", "Primer"), sep = "_") %>%
  mutate(Primer = case_when(Primer == "midV2" ~ "Midnight",
                            Primer == "v5" ~ "Artic")) %>%
  pivot_wider(names_from = c(Primer), values_from = lineage)

datatable(pangolin, options = list(pageLength = -1))
```

